---
title: "R-Ladies Chapter Health Report"
subtitle: "Monitoring Chapter Activity and Engagement"
date: "`r Sys.Date()`"
author: "R-Ladies Global"
slug: chapter-health-report
categories:
  - reports
  - monitoring
  
tags:
  - chapters
  - analytics
draft: false
---

```{r}
#| label: setup
#| include: false
source(here::here("reports/_common.R"))
```

## Executive Summary

This report identifies chapters needing support and highlights successful chapters to help the Chapter Monitoring Team maintain healthy community engagement across the R-Ladies network.

```{r}
#| label: summary-stats
recent_cutoff <- Sys.Date() - months(6)
inactive_cutoff <- Sys.Date() - months(6)

# Parse event dates
events_parsed <- events |>
  mutate(
    event_date = as.Date(date),
    is_active = status %in% c("active", "ACTIVE")
  )

# Calculate metrics
total_chapters <- n_distinct(chapters$urlname)

active_chapters <- events_parsed |>
  filter(
    event_date >= recent_cutoff,
    !status %in% c("cancelled", "CANCELLED")
  ) |>
  distinct(group_urlname) |>
  nrow()

inactive_count <- total_chapters - active_chapters
```

- **Total Chapters**: `r total_chapters`  
- **Active (last 6 months)**: `r active_chapters` (`r round(100 * active_chapters / total_chapters, 1)`%)  
- **Inactive**: `r inactive_count` (`r round(100 * inactive_count / total_chapters, 1)`%)  

## Chapters Needing Immediate Attention

These chapters have not held events in the past 6 months and may need outreach or support.

```{r}
#| label: inactive-chapters
last_activity <- events_parsed |>
  filter(!status %in% c("cancelled", "CANCELLED")) |>
  group_by(group_urlname) |>
  summarise(
    last_event = max(event_date, na.rm = TRUE),
    total_events = n(),
    .groups = "drop"
  )

inactive_chapters <- last_activity |>
  filter(last_event < inactive_cutoff) |>
  arrange(last_event) |>
  mutate(
    days_since = as.numeric(Sys.Date() - last_event),
    months_since = round(days_since / 30, 1),
    Chapter = str_replace_all(group_urlname, "@", " (at) ")
  ) |>
  select(
    Chapter,
    `Last Event` = last_event,
    `Months Since` = months_since,
    `Total Events` = total_events
  )

knitr::kable(
  head(inactive_chapters, 30),
  caption = "Chapters with no activity in 6+ months (showing first 30)"
)
```

**Action Items:**

- Total chapters needing outreach: `r nrow(inactive_chapters)`  
- Chapters inactive 6-12 months: `r sum(inactive_chapters$'Months Since' >= 6 & inactive_chapters$'Months Since' <= 12)`  
- Chapters inactive 12+ months: `r sum(inactive_chapters$'Months Since' > 12)`  

## Most Active Chapters (Benchmarks)

These chapters can serve as models and mentors for others.

```{r}
#| label: active-chapters
#| fig-cap: "Top 20 chapters by events in last 6 months"

active_counts <- events_parsed |>
  filter(
    event_date >= recent_cutoff,
    !status %in% c("cancelled", "CANCELLED")
  ) |>
  count(group_urlname, sort = TRUE)

top_chapters <- active_counts |>
  slice_head(n = 20)

ggplot(top_chapters, aes(x = n, y = reorder(group_urlname, n))) +
  geom_col(fill = rladies_colors$purple) +
  geom_text(
    aes(label = n),
    hjust = -0.2,
    size = 3.5
  ) +
  labs(
    title = "Most Active Chapters (Last 6 Months)",
    x = "Number of Events",
    y = NULL
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.1)))
```

**Benchmark Statistics:**

- Median events per active chapter: `r median(active_counts$n)`  
- Mean events per active chapter: `r round(mean(active_counts$n), 1)`  
- Top performer: `r top_chapters$group_urlname[1]` with `r top_chapters$n[1]` events  

## Activity Trends Over Time

```{r}
#| label: activity-trends
#| fig-cap: "Event activity by month across all chapters"

monthly_activity <- events_parsed |>
  filter(
    !status %in% c("cancelled", "CANCELLED"),
    event_date >= Sys.Date() - months(12)
  ) |>
  mutate(month = floor_date(event_date, "month")) |>
  count(month)

ggplot(monthly_activity, aes(x = month, y = n)) +
  geom_line(color = rladies_colors$purple, linewidth = 1) +
  geom_point(color = rladies_colors$purple, size = 3) +
  geom_smooth(
    method = "loess",
    se = TRUE,
    alpha = 0.2,
    color = rladies_colors$dark_purple
  ) +
  labs(
    title = "Monthly Event Activity (Last 12 Months)",
    x = "Month",
    y = "Number of Events"
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "2 months") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## New vs. Established Chapters

```{r}
#| label: chapter-age
#| fig-cap: "Activity distribution by chapter maturity"

chapter_activity <- events_parsed |>
  filter(!status %in% c("cancelled", "CANCELLED")) |>
  group_by(group_urlname) |>
  summarise(
    first_event = min(event_date, na.rm = TRUE),
    last_event = max(event_date, na.rm = TRUE),
    total_events = n(),
    recent_events = sum(event_date >= recent_cutoff),
    .groups = "drop"
  ) |>
  mutate(
    age_months = as.numeric(Sys.Date() - first_event) / 30,
    maturity = case_when(
      age_months < 6 ~ "New (< 6 months)",
      age_months < 24 ~ "Developing (6-24 months)",
      TRUE ~ "Established (2+ years)"
    ),
    activity_level = case_when(
      recent_events == 0 ~ "Inactive",
      recent_events <= 2 ~ "Low",
      recent_events <= 6 ~ "Moderate",
      TRUE ~ "High"
    )
  )

activity_summary <- chapter_activity |>
  count(maturity, activity_level) |>
  mutate(
    maturity = factor(
      maturity,
      levels = c(
        "New (< 6 months)",
        "Developing (6-24 months)",
        "Established (2+ years)"
      )
    )
  )

ggplot(activity_summary, aes(x = maturity, y = n, fill = activity_level)) +
  geom_col(position = "stack") +
  scale_fill_manual(
    values = c(
      "Inactive" = rladies_colors$grey,
      "Low" = rladies_colors$light_purple,
      "Moderate" = rladies_colors$purple,
      "High" = rladies_colors$dark_purple
    )
  ) +
  labs(
    title = "Chapter Activity by Maturity Level",
    x = "Chapter Maturity",
    y = "Number of Chapters",
    fill = "Activity Level"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Recommendations

### Immediate Actions (Next 30 Days)

1. **Priority Outreach**: Contact the `r sum(inactive_chapters$'Months Since' > 12)` chapters inactive for 12+ months  
2. **Check-in Calls**: Schedule supportive check-ins with chapters inactive 6-12 months  
3. **Success Sharing**: Connect inactive chapters with regional active chapter organizers  

### Medium-term Strategy (Next Quarter)

1. **Mentorship Program**: Pair struggling chapters with successful ones in similar regions/time zones  
2. **Resource Audit**: Ensure all chapter organizers have access to:  
   - Event planning templates  
   - Speaker resources  
   - Marketing materials  
3. **Celebrate Success**: Feature top performing chapters in newsletter  

### Support Needed

- Organizer burnout assessment for long-inactive chapters  
- Regional coordinator support for clusters of inactive chapters  
- New organizer recruitment for chapters with departed leadership  

---

*Report generated: `r format(Sys.time(), "%Y-%m-%d %H:%M %Z")`*  
*Data source: R-Ladies Meetup Archive*