---
title: "R-Ladies Event Analytics"
subtitle: "Understanding Event Patterns and Formats"
date: "`r Sys.Date()`"
author: "R-Ladies Global"
slug: event-analytics
categories:
  - reports
  - analytics
tags:
  - events
  - patterns
  - insights
draft: false
---

```{r}
#| label: setup
#| include: false
source(here::here("reports/_common.R"))

offset_to_hours <- function(offset) {
  offset <- as.character(offset)
  offset[offset == "Z"] <- "+00:00"
  m <- regexec("^([+-])(\\d{2}):?(\\d{2})$", offset)
  parts <- regmatches(offset, m)
  vapply(parts, FUN.VALUE = numeric(1), function(p) {
    if (length(p) == 0) {
      return(NA_real_)
    }
    sign <- if (p[2] == "-") -1L else 1L
    h <- as.integer(p[3])
    mm <- as.integer(p[4])
    as.numeric((sign * (h * 3600L + mm * 60L)) / 3600)
  })
}
```

## Executive Summary

This report analyzes event characteristics across the R-Ladies network to identify successful patterns and guide organizers in planning effective events.

```{r}
#| label: summary-metrics
events_analyzed <- events |>
  filter(!status %in% c("cancelled", "CANCELLED")) |>
  mutate(
    event_date = as.Date(date),
    event_datetime = datetime,
    event_hour = hour(datetime) + offset_to_hours(original_tz),
    event_dow = wday(date, label = TRUE, week_start = 1),
    event_month = month(date, label = TRUE),
    event_tod = case_when(
      event_hour < 8 ~ "Night",
      event_hour < 12 ~ "Morning",
      event_hour < 17 ~ "Afternoon",
      event_hour < 20 ~ "Evening",
      .default = "Night"
    ),
    year = year(date),
    is_virtual = str_detect(
      tolower(location %||% ""),
      "online|virtual|zoom|meet|not announced"
    ),
    duration_hours = as.numeric(duration) / 3600
  )

# Recent events (last 12 months)
recent_events <- events_analyzed |>
  filter(event_date >= Sys.Date() - months(12))

total_events <- nrow(recent_events)
virtual_pct <- mean(recent_events$is_virtual, na.rm = TRUE) * 100
avg_duration <- mean(recent_events$duration_hours, na.rm = TRUE)
```

- **Events analyzed (last 12 months)**: `r format(total_events, big.mark = ",")`  
- **Virtual events**: `r round(virtual_pct, 1)`%  
- **Average event duration**: `r round(avg_duration, 1)` hours  
- **Chapters hosting events**: `r n_distinct(recent_events$group_urlname)`  

## Event Timing Patterns

### Day of Week Distribution

```{r}
#| label: day-of-week
#| fig-cap: "Event distribution by day of week"

dow_counts <- recent_events |>
  filter(!is.na(event_dow)) |>
  count(event_dow) |>
  mutate(percentage = n / sum(n) * 100)

top_dow <- dow_counts |>
  slice_max(n, n = 1)
weekend_total <- dow_counts |>
  filter(event_dow %in% c("Sat", "Sun")) |>
  summarise(n = sum(n), pct = sum(percentage))

ggplot(dow_counts, aes(x = event_dow, y = n)) +
  geom_col(fill = rladies_colors$purple) +
  geom_text(
    aes(label = sprintf("%d\n(%.1f%%)", n, percentage)),
    vjust = -0.5,
    size = 3.5
  ) +
  labs(
    title = "Events by Day of Week (Last 12 Months)",
    x = "Day of Week",
    y = "Number of Events"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

**`r top_dow$event_dow`** is the most popular day with `r top_dow$n` events (`r round(top_dow$percentage, 1)`%).
Weekend events represent `r round(weekend_total$pct, 1)`% of all events.

### Time of Day Patterns

```{r}
#| label: time-of-day
#| fig-cap: "Event start times throughout the day"

hour_counts <- recent_events |>
  filter(!is.na(event_hour)) |>
  count(event_hour, event_tod)

peak_hour <- hour_counts |> slice_max(n, n = 1)
evening_count <- sum(hour_counts$n[
  hour_counts$event_hour >= 17 & hour_counts$event_hour < 20
])
business_count <- sum(hour_counts$n[
  hour_counts$event_hour >= 9 & hour_counts$event_hour < 17
])

ggplot(hour_counts, aes(x = event_hour, y = n)) +
  geom_col(aes(fill = event_tod)) +
  scale_fill_manual(
    values = c(
      "Morning" = rladies_colors$light_purple,
      "Afternoon" = rladies_colors$purple,
      "Evening" = rladies_colors$dark_purple,
      "Night" = rladies_colors$grey
    )
  ) +
  labs(
    title = "Event Start Times (24-hour format)",
    x = "Hour of Day",
    y = "Number of Events",
    fill = "Time Period"
  ) +
  scale_x_continuous(breaks = seq(0, 23, 2))
```

The peak hour is **`r sprintf("%02d:00", peak_hour$event_hour)`** with `r peak_hour$n` events.
Evening events (17:00-20:00): `r evening_count`, business hours (9:00-17:00): `r business_count`.

### Seasonal Trends

```{r}
#| label: seasonal-patterns
#| fig-cap: "Event activity by month"

monthly_patterns <- events_analyzed |>
  filter(year >= year(Sys.Date()) - 2) |>
  count(event_month, year) |>
  group_by(event_month) |>
  summarise(
    avg_events = mean(n),
    .groups = "drop"
  )

busiest_month <- monthly_patterns |> slice_max(avg_events, n = 1)
quietest_month <- monthly_patterns |> slice_min(avg_events, n = 1)

ggplot(monthly_patterns, aes(x = event_month, y = avg_events)) +
  geom_col(fill = rladies_colors$purple) +
  geom_hline(
    yintercept = mean(monthly_patterns$avg_events),
    linetype = "dashed",
    color = rladies_colors$dark_purple
  ) +
  labs(
    title = "Average Events per Month (Last 2 Years)",
    subtitle = "Dashed line shows overall average",
    x = "Month",
    y = "Average Number of Events"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

**`r busiest_month$event_month`** is the busiest month (avg: `r round(busiest_month$avg_events, 1)` events) while **`r quietest_month$event_month`** is the quietest (avg: `r round(quietest_month$avg_events, 1)` events).

## Event Format Analysis

### Virtual vs. In-Person Distribution

```{r}
#| label: venue-format
#| fig-cap: "Event format trends over time"

format_trends <- events_analyzed |>
  filter(year >= year(Sys.Date()) - 2) |>
  mutate(
    quarter = paste0(year, " Q", quarter(event_date)),
    format = if_else(is_virtual, "Virtual", "In-Person")
  ) |>
  count(quarter, format) |>
  group_by(quarter) |>
  mutate(percentage = n / sum(n) * 100)

ggplot(format_trends, aes(x = quarter, y = percentage, fill = format)) +
  geom_col(position = "stack") +
  scale_fill_manual(
    values = c(
      "Virtual" = rladies_colors$purple,
      "In-Person" = rladies_colors$light_purple
    )
  ) +
  labs(
    title = "Event Format Distribution by Quarter",
    x = "Quarter",
    y = "Percentage of Events",
    fill = "Format"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Event Duration Patterns

```{r}
#| label: duration-analysis
#| fig-cap: "Distribution of event durations"

duration_data <- recent_events |>
  filter(!is.na(duration_hours), duration_hours > 0, duration_hours < 12) |>
  mutate(
    duration_category = case_when(
      duration_hours < 1 ~ "< 1 hour",
      duration_hours < 1.5 ~ "1-1.5 hours",
      duration_hours < 2.5 ~ "1.5-2.5 hours",
      duration_hours < 4 ~ "2.5-4 hours",
      TRUE ~ "4+ hours"
    )
  )

duration_summary <- duration_data |>
  count(duration_category) |>
  mutate(
    duration_category = factor(
      duration_category,
      levels = c(
        "< 1 hour",
        "1-1.5 hours",
        "1.5-2.5 hours",
        "2.5-4 hours",
        "4+ hours"
      )
    )
  )

most_common_duration <- duration_summary |> slice_max(n, n = 1)
median_duration <- median(duration_data$duration_hours, na.rm = TRUE)
mean_duration <- mean(duration_data$duration_hours, na.rm = TRUE)

ggplot(duration_summary, aes(x = duration_category, y = n)) +
  geom_col(fill = rladies_colors$purple) +
  geom_text(
    aes(label = sprintf("%d\n(%.1f%%)", n, n / sum(n) * 100)),
    vjust = -0.5,
    size = 3.5
  ) +
  labs(
    title = "Event Duration Distribution",
    x = "Duration",
    y = "Number of Events"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The most common duration is **`r most_common_duration$duration_category`** (`r most_common_duration$n` events, `r round(most_common_duration$n/sum(duration_summary$n)*100, 1)`%).
Median: `r round(median_duration, 1)` hours, mean: `r round(mean_duration, 1)` hours.

## Geographic Distribution

```{r}
#| label: geographic-patterns
#| fig-cap: "Top 20 locations by number of events"

# Extract city from location string (rough heuristic)
location_events <- recent_events |>
  filter(!is.na(location), !is_virtual) |>
  mutate(
    # Try to extract city - it's usually before the last comma or the whole string
    city = str_extract(location, "^[^,]+")
  ) |>
  filter(!is.na(city)) |>
  count(city, sort = TRUE) |>
  slice_head(n = 20)

ggplot(location_events, aes(x = n, y = reorder(city, n))) +
  geom_col(fill = rladies_colors$purple) +
  geom_text(
    aes(label = n),
    hjust = -0.2,
    size = 3
  ) +
  labs(
    title = "Most Active Locations (Last 12 Months)",
    x = "Number of Events",
    y = NULL
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.1)))
```

## Event Type Distribution

```{r}
#| label: event-type
#| fig-cap: "Event status distribution over time"

type_trends <- events_analyzed |>
  filter(year >= year(Sys.Date()) - 2) |>
  mutate(quarter = paste0(year, " Q", quarter(event_date))) |>
  count(quarter, is_virtual) |>
  group_by(quarter) |>
  mutate(percentage = n / sum(n) * 100)

ggplot(
  type_trends,
  aes(x = quarter, y = n, fill = is_virtual)
) +
  geom_col(position = "stack") +
  labs(
    title = "Event Type by Quarter",
    x = "Quarter",
    y = "Number of Events",
    fill = "Virtual Event"
  ) +
  scale_fill_manual(
    values = c(
      "TRUE" = rladies_colors$purple,
      "FALSE" = rladies_colors$light_purple
    ),
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Recommendations for Organizers

### Timing Optimization

```{r}
#| label: timing-recommendations
#| include: false

# Calculate recommendations dynamically
top_3_days <- dow_counts |>
  slice_max(n, n = 3) |>
  pull(event_dow) |>
  as.character()

top_3_hours <- hour_counts |>
  slice_max(n, n = 3) |>
  pull(event_hour)
```

- Schedule events on **`r paste(top_3_days, collapse = ", ")`** for highest participation  
- Consider **`r sprintf("%02d:00-%02d:00", min(top_3_hours), max(top_3_hours))`** for optimal timing  
- Plan for reduced activity in **`r quietest_month$event_month`**, ramp up in **`r busiest_month$event_month`**  

### Format Selection

- **Virtual events**: Greater accessibility, easier speaker recruitment, weather-independent  
- **In-person events**: Stronger networking, hands-on workshops, community building  
- **Hybrid approach**: Consider alternating formats based on content type  

### Duration Guidelines

Based on observed patterns:

- **Standard meetups**: `r most_common_duration$duration_category` (most popular)  
- **Workshops**: 2.5-4 hours for hands-on learning  
- **Social events**: 1-1.5 hours to encourage participation  

### Event Series Ideas

Based on successful patterns:

- **Regular meetups**: Consistent schedule on popular days  
- **Quarterly workshops**: Longer format for skill-building  
- **Annual celebrations**: Large events for milestones (IWD, R-Ladies Day)  
- **Co-working sessions**: Regular low-structure community time  

---

*Report generated: `r format(Sys.time(), "%Y-%m-%d %H:%M %Z")`*  
*Data source: R-Ladies Meetup Archive*

