---
title: "R-Ladies Annual Impact Report"
subtitle: "Community Growth and Achievements"
date: "`r Sys.Date()`"
author: "R-Ladies Global"
slug: funder-report
categories:
  - reports
  - annual
tags:
  - impact
  - funders
  - metrics
draft: false
---

```{r}
#| label: setup
#| include: false
source(here::here("reports/_common.R"))
library(maps)
```

## Geographic Analysis

This report maps the global footprint of R-Ladies, showing where chapters are located and how they cluster across countries and regions.

### Active countries

```{r}
#| label: active-chapters
#| include: false

# Restrict to chapters active in the last year, working from `events` just like
# funder-report.qmd and chapter-health.qmd. chapter_country / chapter_country_iso
# are joined onto events in reports/_common.R (the chapter's country, not the
# venue's), so geography comes straight from the event table.
report_end <- Sys.Date()
report_start <- report_end - years(1)

active_chapters <- events |>
  mutate(event_date = as.Date(datetime_utc)) |>
  filter(
    event_date >= report_start,
    event_date < report_end,
    !status %in% c("cancelled", "CANCELLED")
  ) |>
  # one row per active chapter, carrying its country fields
  distinct(group_urlname, chapter_country, chapter_country_iso)

total_chapters <- n_distinct(active_chapters$group_urlname)
total_countries <- n_distinct(active_chapters$chapter_country_iso)
```

In the last year, **`r total_chapters`** R-Ladies chapters were active across **`r total_countries`** countries worldwide.

## Chapters by Country

```{r}
#| label: choropleth-map
#| fig-cap: "Number of active R-Ladies chapters per country (last year)"
#| fig-width: 10
#| fig-height: 5.5

# Count active chapters per country. Reconcile the few country names that the
# Meetup data spells differently from maps::map_data("world").
chapters_by_country <- active_chapters |>
  count(chapter_country, name = "n_chapters") |>
  mutate(
    map_region = recode(
      chapter_country,
      "United States of America" = "USA",
      "United Kingdom of Great Britain and Northern Ireland" = "UK",
      "Iran (Islamic Republic of)" = "Iran",
      "Bolivia (Plurinational State of)" = "Bolivia",
      "Venezuela (Bolivarian Republic of)" = "Venezuela",
      "Viet Nam" = "Vietnam",
      "Türkiye" = "Turkey"
    )
  )

# World basemap from the `maps` package (bundled with ggplot2), joined to the
# chapter counts by country name.
world_chapters <- map_data("world") |>
  filter(region != "Antarctica") |>
  left_join(
    chapters_by_country |> select(map_region, n_chapters),
    by = c("region" = "map_region")
  )

# Discretize chapter counts into labelled ranges for a categorical legend.
world_chapters <- world_chapters |>
  mutate(
    chapter_bin = cut(
      n_chapters,
      breaks = c(0, 1, 2, 3, 5, Inf),
      labels = c("1", "2", "3", "4–5", "6+"),
      right = TRUE
    ),
    chapter_bin = addNA(chapter_bin)
  )
levels(world_chapters$chapter_bin)[is.na(levels(world_chapters$chapter_bin))] <- "none"

ggplot(
  world_chapters,
  aes(x = long, y = lat, group = group, fill = chapter_bin)
) +
  geom_polygon(color = "white", linewidth = 0.1) +
  scale_fill_manual(
    values = c(
      "1"    = rladies_colors$light_purple,
      "2"    = "#E07AB8",
      "3"    = rladies_colors$purple,
      "4–5"  = "#6E2E6F",
      "6+"   = rladies_colors$dark_purple,
      "none" = rladies_colors$grey
    ),
    breaks = c("1", "2", "3", "4–5", "6+"),
    name = "Chapters per country",
    guide = guide_legend(
      title.position = "top",
      title.hjust = 0.5,
      keywidth = 1.4,
      keyheight = 0.8,
      nrow = 1,
      label.position = "bottom"
    )
  ) +
  coord_fixed(ratio = 1.3, xlim = c(-180, 180)) +
  labs(
    title = "Active R-Ladies Chapters Around the World",
    subtitle = paste(
      total_chapters, "active chapters across", total_countries, "countries",
      "(last 12 months)"
    ),
    x = NULL,
    y = NULL
  ) +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(size = 9, color = rladies_colors$dark_purple),
    legend.text = element_text(size = 8)
  )
```

Countries shown in grey have no active R-Ladies chapter; darker purple indicates a greater number of active chapters in that country.

## Top Countries by Chapter Count

```{r}
#| label: top-countries
#| fig-cap: "Countries with the most active R-Ladies chapters"

top_countries <- chapters_by_country |>
  slice_max(n_chapters, n = 15) |>
  mutate(chapter_country = reorder(chapter_country, n_chapters))

ggplot(top_countries, aes(x = n_chapters, y = chapter_country)) +
  geom_col(fill = rladies_colors$purple) +
  geom_text(
    aes(label = n_chapters),
    hjust = -0.2,
    size = 3.5
  ) +
  labs(
    title = "Top Countries by Number of Active Chapters",
    x = "Number of Chapters",
    y = NULL
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.12)))
```

## Chapters by Region

```{r}
#| label: chapters-by-region
#| fig-cap: "Active chapter distribution across the six reporting regions"

# Bucket active chapters by region using the chapter's ISO2 code, exactly as
# the event reports do via chapter_region() from reports/_common.R.
chapters_by_region <- active_chapters |>
  mutate(
    region = chapter_region(chapter_country_iso),
    region = tidyr::replace_na(region, "Other")
  ) |>
  count(region, sort = TRUE)

ggplot(chapters_by_region, aes(x = n, y = reorder(region, n))) +
  geom_col(fill = rladies_colors$purple) +
  geom_text(
    aes(label = sprintf("%d (%.1f%%)", n, n / sum(n) * 100)),
    hjust = -0.1,
    size = 3.5
  ) +
  labs(
    title = "Active Chapters by Region",
    subtitle = "Chapters active in the last year, grouped into the six regions",
    x = "Number of Chapters",
    y = NULL
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15)))
```

```{r}
#| label: region-metrics
#| include: false

regions_with_chapters <- chapters_by_region |>
  filter(region != "Other") |>
  nrow()

leading_region <- chapters_by_region |>
  slice_max(n, n = 1) |>
  pull(region)
```

Active chapters are present in **`r regions_with_chapters`** of the six reporting regions, with **`r leading_region`** hosting the largest share.

### Growing countries

```{r}
#| label: growing-countries
#| fig-cap: "Top 5 countries by year-over-year attendance growth (last 12 months)"

# Attendance ("going") by chapter country, comparing the last 12 months against
# the 12 months before that. 

previous_start <- report_start - years(1)

attendance_periods <- events |>
  mutate(event_date = as.Date(datetime_utc)) |>
  filter(!status %in% c("cancelled", "CANCELLED")) |>
  mutate(
    period = case_when(
      event_date >= report_start & event_date < report_end ~ "current",
      event_date >= previous_start & event_date < report_start ~ "previous",
      TRUE ~ NA_character_
    )
  ) |>
  filter(!is.na(period)) |>
  group_by(chapter_country, period) |>
  summarise(attendance = sum(going, na.rm = TRUE), .groups = "drop") |>
  tidyr::pivot_wider(names_from = period, values_from = attendance, values_fill = 0)

# Keep countries present in both periods, compute YoY % growth, take top 5 of only countries that have shown growth.
growing_countries <- attendance_periods |>
  filter(previous > 0, current > previous) |>
  mutate(growth_pct = (current - previous) / previous * 100) |>
  slice_max(growth_pct, n = 5) |>
  mutate(chapter_country = reorder(chapter_country, growth_pct))

ggplot(growing_countries, aes(x = growth_pct, y = chapter_country)) +
  geom_col(fill = rladies_colors$purple) +
  geom_text(
    aes(label = sprintf("+%.0f%% (%d→%d)", growth_pct, previous, current)),
    hjust = -0.05,
    size = 3.5
  ) +
  labs(
    title = "Highest Growing Countries by Attendance",
    subtitle = "Year-over-year change in total attendance (last 12 months)",
    x = "Attendance Growth (%)",
    y = NULL
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.25)))
```


### Event type by country and region

```{r}
#| label: event-type-base
#| fig-cap: "In-person vs online events by region (last 12 months)"

# Classify last-year events as Online vs In-Person using the same location
# heuristic as funder-report.qmd, and attach the chapter's country and region
# (chapter_country / chapter_country_iso are joined onto events in _common.R).
events_typed <- events |>
  mutate(event_date = as.Date(datetime_utc)) |>
  filter(
    event_date >= report_start,
    event_date < report_end,
    !status %in% c("cancelled", "CANCELLED")
  ) |>
  mutate(
    format = if_else(
      str_detect(tolower(location %||% ""), "online|virtual|zoom|meet|not announced"),
      "Online",
      "In-Person"
    ),
    region = tidyr::replace_na(chapter_region(chapter_country_iso), "Other")
  )

# Count by region and format.
type_by_region <- events_typed |>
  count(region, format)

ggplot(type_by_region, aes(x = n, y = reorder(region, n), fill = format)) +
  geom_col() +
  scale_fill_manual(
    values = c(
      "Online" = rladies_colors$purple,
      "In-Person" = rladies_colors$light_purple
    )
  ) +
  labs(
    title = "Event Format by Region",
    subtitle = "In-person and online events (last 12 months)",
    x = "Number of Events",
    y = NULL,
    fill = "Format"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)))
```

```{r}
#| label: event-type-country
#| fig-cap: "In-person vs online events by country (last 12 months)"
#| fig-height: 7

# Same breakdown at country level. Many countries have few events, so we show
# the top 20 by total events to keep the axis readable.
type_by_country <- events_typed |>
  count(chapter_country, format) |>
  group_by(chapter_country) |>
  mutate(total = sum(n)) |>
  ungroup() |>
  slice_max(total, n = 40, with_ties = FALSE) |>  # 40 rows = up to 20 countries x 2 formats
  mutate(chapter_country = reorder(chapter_country, total))

ggplot(type_by_country, aes(x = n, y = chapter_country, fill = format)) +
  geom_col() +
  scale_fill_manual(
    values = c(
      "Online" = rladies_colors$purple,
      "In-Person" = rladies_colors$light_purple
    )
  ) +
  labs(
    title = "Event Format by Country",
    subtitle = "Top countries by total events (last 12 months)",
    x = "Number of Events",
    y = NULL,
    fill = "Format"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)))
```

```{r}
#| label: event-type-trend-global
#| fig-cap: "Monthly trend of event formats (last 12 months)"

# Monthly counts of Online vs In-Person across all chapters.
trend_global <- events_typed |>
  mutate(month = lubridate::floor_date(event_date, "month")) |>
  count(month, format)

ggplot(trend_global, aes(x = month, y = n, color = format)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_color_manual(
    values = c(
      "Online" = rladies_colors$purple,
      "In-Person" = rladies_colors$light_purple
    )
  ) +
  labs(
    title = "Monthly Event Format Trend",
    subtitle = "Last 12 months, all chapters",
    x = "Month",
    y = "Number of Events",
    color = "Format"
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "2 months") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| label: event-type-trend-region
#| fig-cap: "Monthly trend of event formats by region (last 12 months)"
#| fig-height: 7

# Same monthly trend, faceted by region. "Other" is dropped to avoid a sparse,
# uninformative panel; remove the filter to keep it.
trend_region <- events_typed |>
  filter(region != "Other") |>
  mutate(month = lubridate::floor_date(event_date, "month")) |>
  count(month, region, format)

ggplot(trend_region, aes(x = month, y = n, color = format)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  facet_wrap(~ region, ncol = 2) +
  scale_color_manual(
    values = c(
      "Online" = rladies_colors$purple,
      "In-Person" = rladies_colors$light_purple
    )
  ) +
  labs(
    title = "Monthly Event Format Trend by Region",
    subtitle = "Last 12 months",
    x = "Month",
    y = "Number of Events",
    color = "Format"
  ) +
  scale_x_date(date_labels = "%b", date_breaks = "3 months") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Venue diversity by chapter

```{r}
#| label: venue-diversity-calc
#| include: false

# Count distinct venues per chapter over the last 12 months. A venue is either
# a distinct in-person venue_name OR "Online" treated as a single venue type.
# Example: 5 events = 2 online + 3 at 3 different places -> 4 venues.
venue_diversity <- events |>
  mutate(event_date = as.Date(datetime_utc)) |>
  filter(
    event_date >= report_start,
    event_date < report_end,
    !status %in% c("cancelled", "CANCELLED")
  ) |>
  mutate(
    # Collapse all online events into one venue label; keep in-person venues
    # as their own names. Unnamed in-person events fall back to "Unknown venue".
    venue = case_when(
      str_detect(tolower(location %||% ""), "online|virtual|zoom|meet|not announced") ~ "Online",
      !is.na(venue_name) & venue_name != "" ~ venue_name,
      TRUE ~ "Unknown venue"
    )
  ) |>
  group_by(group_urlname, group_name) |>
  summarise(n_venues = n_distinct(venue), .groups = "drop") |>
  arrange(desc(n_venues))

# Headline indicator: the single most venue-diverse chapter.
most_diverse_chapter <- venue_diversity |> slice_max(n_venues, n = 1, with_ties = FALSE)
median_venues <- median(venue_diversity$n_venues)
```

The most venue-diverse chapter in the last year is
**`r most_diverse_chapter$group_name`**, with
**`r most_diverse_chapter$n_venues`** distinct venues (online counts as one).
The median active chapter used **`r median_venues`** venue(s).

```{r}
#| label: venue-diversity-plot
#| fig-cap: "Chapters with the most venue variety (last 12 months)"

# Top 15 chapters by number of distinct venues. group_name is the readable
# chapter label; fall back to group_urlname only if a name is missing.
top_venue_chapters <- venue_diversity |>
  slice_max(n_venues, n = 15, with_ties = FALSE) |>
  mutate(
    chapter = coalesce(group_name, group_urlname),
    chapter = reorder(chapter, n_venues)
  )

ggplot(top_venue_chapters, aes(x = n_venues, y = chapter)) +
  geom_col(fill = rladies_colors$purple) +
  geom_text(
    aes(label = n_venues),
    hjust = -0.3,
    size = 3.5
  ) +
  labs(
    title = "Most Venue-Diverse Chapters",
    subtitle = "Distinct venues per chapter, online counted as one (last 12 months)",
    x = "Number of Distinct Venues",
    y = NULL
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.12)), breaks = scales::breaks_pretty())
```

---

*Report period: `r format(report_start, "%B %d, %Y")` - `r format(report_end, "%B %d, %Y")`*  
*Report generated: `r format(Sys.time(), "%Y-%m-%d %H:%M %Z")`*  
*Data source: R-Ladies Meetup Archive*
