---
title: "R-Ladies Quarterly Summary"
subtitle: "Period-over-Period Community Insights"
date: "`r Sys.Date()`"
author: "R-Ladies Global"
slug: quarterly-summary
categories:
  - reports
  - quarterly
tags:
  - summary
  - trends
  - metrics
draft: false
---

```{r}
#| label: setup
#| include: false
source(here::here("reports/_common.R"))
```

## Executive Summary

This quarterly report provides a snapshot of R-Ladies community activity, highlighting key metrics and trends compared to previous periods.

```{r}
#| label: quarterly-metrics
# Define current and previous quarters
current_quarter_end <- floor_date(Sys.Date(), "quarter") + months(3) - days(1)
current_quarter_start <- floor_date(current_quarter_end, "quarter")

previous_quarter_end <- current_quarter_start - days(1)
previous_quarter_start <- floor_date(previous_quarter_end, "quarter")

year_ago_quarter_end <- current_quarter_end - years(1)
year_ago_quarter_start <- current_quarter_start - years(1)

# Prepare events data
events_analyzed <- events |>
  filter(!status %in% c("cancelled", "CANCELLED")) |>
  mutate(
    event_date = as.Date(datetime_utc),
    is_virtual = str_detect(
      tolower(location %||% ""),
      "online|virtual|zoom|meet|not announced"
    )
  )

# Current quarter
current_events <- events_analyzed |>
  filter(event_date >= current_quarter_start, event_date <= current_quarter_end)

# Previous quarter
previous_events <- events_analyzed |>
  filter(
    event_date >= previous_quarter_start,
    event_date <= previous_quarter_end
  )

# Year ago quarter
year_ago_events <- events_analyzed |>
  filter(
    event_date >= year_ago_quarter_start,
    event_date <= year_ago_quarter_end
  )

# Calculate metrics
current_total <- nrow(current_events)
previous_total <- nrow(previous_events)
year_ago_total <- nrow(year_ago_events)

current_chapters <- n_distinct(current_events$group_urlname)
previous_chapters <- n_distinct(previous_events$group_urlname)
year_ago_chapters <- n_distinct(year_ago_events$group_urlname)

# Growth rates
qoq_growth <- round((current_total - previous_total) / previous_total * 100, 1)
yoy_growth <- round((current_total - year_ago_total) / year_ago_total * 100, 1)

chapter_qoq <- round(
  (current_chapters - previous_chapters) / previous_chapters * 100,
  1
)
chapter_yoy <- round(
  (current_chapters - year_ago_chapters) / year_ago_chapters * 100,
  1
)

# Format quarter labels
current_q <- paste0(
  year(current_quarter_start),
  " Q",
  quarter(current_quarter_start)
)
previous_q <- paste0(
  year(previous_quarter_start),
  " Q",
  quarter(previous_quarter_start)
)
year_ago_q <- paste0(
  year(year_ago_quarter_start),
  " Q",
  quarter(year_ago_quarter_start)
)
```

### Quarter: `r current_q`

**Period**: `r format(current_quarter_start, "%B %d")` - `r format(current_quarter_end, "%B %d, %Y")`

- **Total Events**: `r format(current_total, big.mark = ",")`  
  - Quarter-over-quarter: `r if_else(qoq_growth >= 0, paste0("+", qoq_growth), as.character(qoq_growth))`%  
  - Year-over-year: `r if_else(yoy_growth >= 0, paste0("+", yoy_growth), as.character(yoy_growth))`%  
- **Active Chapters**: `r current_chapters`  
  - Quarter-over-quarter: `r if_else(chapter_qoq >= 0, paste0("+", chapter_qoq), as.character(chapter_qoq))`%  
  - Year-over-year: `r if_else(chapter_yoy >= 0, paste0("+", chapter_yoy), as.character(chapter_yoy))`%  

## Quarterly Comparison

### Event Volume Trends

```{r}
#| label: quarterly-comparison
#| fig-cap: "Event volume comparison across quarters"

quarterly_data <- tibble(
  quarter = c(year_ago_q, previous_q, current_q),
  events = c(year_ago_total, previous_total, current_total),
  quarter_date = c(
    year_ago_quarter_start,
    previous_quarter_start,
    current_quarter_start
  )
) |>
  mutate(
    quarter = factor(quarter, levels = c(year_ago_q, previous_q, current_q))
  )

ggplot(quarterly_data, aes(x = quarter, y = events)) +
  geom_col(fill = rladies_colors$purple) +
  geom_text(
    aes(label = format(events, big.mark = ",")),
    vjust = -0.5,
    size = 4
  ) +
  labs(
    title = "Event Volume Comparison",
    subtitle = "Last three quarters",
    x = "Quarter",
    y = "Number of Events"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

### Active Chapter Trends

```{r}
#| label: chapter-comparison
#| fig-cap: "Active chapter comparison across quarters"

chapter_quarterly <- tibble(
  quarter = c(year_ago_q, previous_q, current_q),
  chapters = c(year_ago_chapters, previous_chapters, current_chapters)
) |>
  mutate(
    quarter = factor(quarter, levels = c(year_ago_q, previous_q, current_q))
  )

ggplot(chapter_quarterly, aes(x = quarter, y = chapters)) +
  geom_col(fill = rladies_colors$purple) +
  geom_text(
    aes(label = chapters),
    vjust = -0.5,
    size = 4
  ) +
  labs(
    title = "Active Chapters Comparison",
    subtitle = "Chapters hosting at least one event per quarter",
    x = "Quarter",
    y = "Number of Active Chapters"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

## Current Quarter Highlights

### Event Format Distribution

```{r}
#| label: format-distribution
#| fig-cap: "Virtual vs in-person events this quarter"

format_data <- current_events |>
  mutate(format = if_else(is_virtual, "Virtual", "In-Person")) |>
  count(format) |>
  mutate(percentage = n / sum(n) * 100)

ggplot(format_data, aes(x = "", y = n, fill = format)) +
  geom_col(width = 1) +
  coord_polar("y") +
  scale_fill_manual(
    values = c(
      "Virtual" = rladies_colors$purple,
      "In-Person" = rladies_colors$light_purple
    )
  ) +
  geom_text(
    aes(label = sprintf("%d\n(%.1f%%)", n, percentage)),
    position = position_stack(vjust = 0.5),
    size = 4
  ) +
  labs(
    title = paste("Event Format Distribution:", current_q),
    fill = "Format"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    legend.position = "bottom"
  )
```

### Monthly Activity Within Quarter

```{r}
#| label: monthly-activity
#| fig-cap: "Event activity by month within current quarter"

monthly_current <- current_events |>
  mutate(month = floor_date(event_date, "month")) |>
  count(month)

ggplot(monthly_current, aes(x = month, y = n)) +
  geom_col(fill = rladies_colors$purple) +
  geom_text(
    aes(label = n),
    vjust = -0.5,
    size = 4
  ) +
  labs(
    title = paste("Monthly Event Distribution:", current_q),
    x = "Month",
    y = "Number of Events"
  ) +
  scale_x_date(date_labels = "%B", date_breaks = "1 month") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

### Top Performing Chapters

```{r}
#| label: top-chapters
#| fig-cap: "Most active chapters this quarter"

top_current <- current_events |>
  count(group_urlname, sort = TRUE) |>
  slice_head(n = 15)

ggplot(top_current, aes(x = n, y = reorder(group_urlname, n))) +
  geom_col(fill = rladies_colors$purple) +
  geom_text(
    aes(label = n),
    hjust = -0.2,
    size = 3.5
  ) +
  labs(
    title = paste("Top 15 Chapters:", current_q),
    x = "Number of Events",
    y = NULL
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15)))
```

## Regional Performance

```{r}
#| label: regional-performance
#| fig-cap: "Event distribution by region this quarter"

regional_current <- current_events |>
  mutate(
    country_code = str_extract(location, "[A-Z]{2}$"),
    region = case_when(
      country_code %in% c("US", "CA", "MX") ~ "North America",
      country_code %in% c("BR", "AR", "CL", "CO", "PE", "UY") ~ "Latin America",
      country_code %in%
        c("GB", "FR", "DE", "ES", "IT", "NL", "BE", "CH", "AT") ~ "Europe",
      country_code %in% c("AU", "NZ") ~ "Oceania",
      country_code %in% c("CN", "JP", "KR", "IN", "SG") ~ "Asia",
      country_code %in% c("ZA", "KE", "NG") ~ "Africa",
      is_virtual ~ "Virtual (Global)",
      TRUE ~ "Other"
    )
  ) |>
  filter(region != "Other") |>
  count(region, sort = TRUE) |>
  mutate(percentage = n / sum(n) * 100)

ggplot(regional_current, aes(x = n, y = reorder(region, n))) +
  geom_col(fill = rladies_colors$purple) +
  geom_text(
    aes(label = sprintf("%d (%.1f%%)", n, percentage)),
    hjust = -0.1,
    size = 3.5
  ) +
  labs(
    title = paste("Regional Distribution:", current_q),
    x = "Number of Events",
    y = NULL
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.2)))
```

## New and Returning Chapters

```{r}
#| label: chapter-status
#| fig-cap: "Chapter activity status"

# Identify new vs returning chapters
new_chapters <- current_events |>
  anti_join(
    previous_events |> select(group_urlname),
    by = "group_urlname"
  ) |>
  distinct(group_urlname)

returning_chapters <- current_events |>
  semi_join(
    previous_events |> select(group_urlname),
    by = "group_urlname"
  ) |>
  distinct(group_urlname)

inactive_chapters <- previous_events |>
  anti_join(
    current_events |> select(group_urlname),
    by = "group_urlname"
  ) |>
  distinct(group_urlname)

chapter_status <- tibble(
  status = c("New/Reactivated", "Continuing", "Now Inactive"),
  count = c(
    nrow(new_chapters),
    nrow(returning_chapters),
    nrow(inactive_chapters)
  )
)

ggplot(chapter_status, aes(x = status, y = count)) +
  geom_col(fill = rladies_colors$purple) +
  geom_text(
    aes(label = count),
    vjust = -0.5,
    size = 4
  ) +
  labs(
    title = "Chapter Activity Status",
    subtitle = paste("Comparing", current_q, "to", previous_q),
    x = NULL,
    y = "Number of Chapters"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

**New or reactivated chapters** (`r nrow(new_chapters)` total):
`r if(nrow(new_chapters) > 0 && nrow(new_chapters) <= 20) paste("-", new_chapters$group_urlname, collapse = "\n") else if(nrow(new_chapters) > 20) paste("See full list in appendix (", nrow(new_chapters), "chapters)") else "None this quarter"`

## Event Timing Patterns

```{r}
#| label: timing-patterns
#| fig-cap: "Day of week distribution this quarter"

dow_current <- current_events |>
  mutate(
    event_datetime = ymd_hms(start, quiet = TRUE),
    dow = wday(event_date, label = TRUE, week_start = 1)
  ) |>
  filter(!is.na(dow)) |>
  count(dow) |>
  mutate(percentage = n / sum(n) * 100)

ggplot(dow_current, aes(x = dow, y = n)) +
  geom_col(fill = rladies_colors$purple) +
  geom_text(
    aes(label = sprintf("%d\n(%.1f%%)", n, percentage)),
    vjust = -0.3,
    size = 3.5
  ) +
  labs(
    title = paste("Events by Day of Week:", current_q),
    x = "Day of Week",
    y = "Number of Events"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

## Performance Indicators

```{r}
#| label: performance-table

# Calculate key metrics for comparison
metrics_comparison <- tibble(
  Metric = c(
    "Total Events",
    "Active Chapters",
    "Avg Events/Chapter",
    "Virtual Events (%)",
    "In-Person Events (%)"
  ),
  !!sym(current_q) := c(
    current_total,
    current_chapters,
    round(current_total / current_chapters, 1),
    round(
      sum(current_events$is_virtual, na.rm = TRUE) / current_total * 100,
      1
    ),
    round(
      sum(!current_events$is_virtual, na.rm = TRUE) / current_total * 100,
      1
    )
  ),
  !!sym(previous_q) := c(
    previous_total,
    previous_chapters,
    round(previous_total / previous_chapters, 1),
    round(
      sum(previous_events$is_virtual, na.rm = TRUE) / previous_total * 100,
      1
    ),
    round(
      sum(!previous_events$is_virtual, na.rm = TRUE) / previous_total * 100,
      1
    )
  ),
  !!sym(year_ago_q) := c(
    year_ago_total,
    year_ago_chapters,
    round(year_ago_total / year_ago_chapters, 1),
    round(
      sum(year_ago_events$is_virtual, na.rm = TRUE) / year_ago_total * 100,
      1
    ),
    round(
      sum(!year_ago_events$is_virtual, na.rm = TRUE) / year_ago_total * 100,
      1
    )
  )
)

knitr::kable(
  metrics_comparison,
  caption = "Key metrics comparison across quarters"
)
```

## Looking Ahead

### Upcoming Quarter Preparation

Based on current trends and historical patterns:

- Expected events next quarter: **`r round(current_total * 1.05)`** (assuming 5% growth)  
- Chapters needing support: **`r nrow(inactive_chapters)`** (became inactive this quarter)  
- Growth opportunity: Focus on `r if(nrow(new_chapters) > 0) paste("sustaining momentum of", nrow(new_chapters), "new/reactivated chapters") else "reactivating dormant chapters"`  

### Recommendations

1. **Chapter support**: Reach out to `r nrow(inactive_chapters)` chapters that went inactive  
2. **New chapter onboarding**: Provide resources to `r nrow(new_chapters)` new/reactivated chapters  
3. **Best practices**: Share success strategies from top performing chapters  
4. **Regional focus**: `r regional_current |> slice_min(n, n = 1) |> pull(region)` shows lowest activity - consider targeted support  

---

*Report period: `r format(current_quarter_start, "%B %d")` - `r format(current_quarter_end, "%B %d, %Y")`*  
*Report generated: `r format(Sys.time(), "%Y-%m-%d %H:%M %Z")`*  
*Data source: R-Ladies Meetup Archive*
